---
title: "Crash Data Pull Template"
output:
  html_document:
    df_print: paged
  word_document: default
---

This is a template to do a data pull. You must select the years of the data.  This notebook uses functions from an R script (Crash_Analysis_Functions.R) to import data (old and new db), to add columns of crash flags (i.e. speed, deer) and to add other columns (i.e. crash time, age group, county name).

## Here are the functions in Crash_Analysis_Functions.R

### Import data functions for new db:
+ import_crashes(fileloc = file_loc, years_selected = years)         - also for vehicle and persons  

### Import data functions for old db:
+ import_crashes_old(fileloc = file_loc, years_selected = years_old) - also for vehicle and persons  
+ These also renames columns (and some variables) "ACCDNMBR", "ACCDDATE", "ACCDMTH", "ACCDSVR", "ACCDHOUR", "INJSVR", "ACCDTYPE" to match new db  

### Crash flag functions add a flag column with either 'Y' or 'N'. These functions and their flag column are:
+ get_deer_crashes(crash_df)            - deer_flag (old & new db)  
+ get_distracted_driver_flag(person_df) - distracted_flag (new db)  
+ get_impaired_driver(person_df)        - impaireddriver_flag (new db)  
+ get_speed_flag(person_df)             - speed_flag (old & new db)  
+ get_teen_driver(person_df)            - teendriver_flag (new db)  
+ get_older_driver(person_df)           - olderdriver_flag (new db)  
+ get_seatbelt_flag_by_role(person_df)  - seatbelt_flag (new db)    # Note: someone in that unit & role did not wear a seatbelt, i.e. a passenger or a driver  
+ get_seatbelt_flag_by_unit(person_df)  - seatbelt_flag (new db)    # Note: someone in that unit did not wear a seatbelt  

### Functions that add a column (only for new db):
+ get_crash_times(any_df)               - newtime  
+ get_age_groups(person_df)             - age_group  
+ county_rename(any_df)                 - countyname  
+ bin_injury_persons(person_df)         - inj - bins into Killed, Injured, No Injury  
+ get_drug_alc_suspected                - drug_alc_susp

### Other functions:
+ get_motorcycle_persons(person_df, vehicle_df) - selects motorcyclists (old & new db)  

## There also snippets which are chunks of code (Data_Pull_Snippets.snippets).

### These include:
+ my_barchart_inj_killed_by_year  
+ my_barchart_mostharm_by_crshsvr_percent  
+ my_stacked_barchart_suspected_drugalcohol  
+ my_barchart_by_hwyclass  
+ my_heatmap_time_of_day  
+ my_barchart_stacked_age_gender  
+ my_make_freq_table_expss  
+ my_save_as_excel_workbook  
+ my_make_county_prop_symbol_map  
+ my_load_county_map_data  

```{r Setup, include = FALSE}
# Set working directory to where this notebook is saved
knitr::opts_chunk$set(echo = FALSE, comment = NA, warning = FALSE, error = FALSE, message = FALSE, tidy = TRUE, fig.dim = c(7, 4), global.par = TRUE)
knitr::opts_knit$set(root.dir = '/...')
```

```{r Load Libraries}
# setwd("W:/HSSA/Keep/Jaclyn Ziebert/R/CSV") # if you want to change wd, put this in the code block
# install.packages("tidyverse", repo = 'https://cran.R-project.org')  try this to install packages in base R, not RStudio

library(dplyr)        # to subset, filter, summarize data, in tidyverse, replaces plyr
library(ggplot2)      # create charts
library(tidyverse)    # to select, filter data, for data science
# library(data.table) # Data type
library(lubridate)    # for working with dates, need this for data import
# library(janitor)    # make tabulation tables, tabyl
library(expss)        # SPSS style package
library(fst)          # Data type of db data
library(ggtext)       # HTML for charts

source("W:/HSSA/Keep/Jaclyn Ziebert/R/Crash-Data-Pulls/Crash_Analysis_Functions.R") # Source of all the functions
```

```{r What to select}
file_loc = "C:/CSV/csv_from_sas/fst/" # location of FST crash database files to load
years = c("17", "18", "19") # select years to import, must be in this format
years_old = c( "16") # select years to import, 2016 and before
```

```{r Run import and combine old/new data}
# This imports new databases - you may choose to select certain columns
all_crashes <- import_crashes() #%>% select(CRSHNMBR, CRSHDATE, CNTYCODE, CRSHSVR, CRSHTYPE, starts_with("ANMLTY")) # columns to select
all_vehicles <- import_vehicles() #%>% select(CRSHNMBR, CRSHDATE, UNITNMBR, VEHTYPE)
all_persons <- import_persons() #%>% select(CRSHNMBR, CNTYCODE, CRSHDATE, UNITNMBR, DRVRFLAG, HLMTUSE, EYEPROT, SFTYEQP, ROLE, SEX, AGE, DISTACT, starts_with("DRVRPC"), starts_with("STATNM"), starts_with("DRVRDS"))

# This imports old databases, does remove parking lot crashes and relabels columns
all_crashes_old <- import_crashes_old()
all_vehicles_old <- import_vehicles_old()
all_persons_old <- import_persons_old()

# To combine old and new db
# crashes <- bind_cols(all_crashes_old, all_crashes)
# vehicles <- bind_cols(all_vehicles_old, all_vehicles)
# persons <- bind_cols(all_persons_old, all_persons)
```

```{r Specific stuff}
# remove duplicates
# motorcycles[!duplicated(motorcycles[, c('CRSHNMBR')]),]

# DRVRPC - put data in long format to count frequency
# all_persons %>% select(starts_with("DRVRPC")) %>% pivot_longer(starts_with("DRVRPC")) %>% filter(value !='') 
```
```{r Create a Table to Knit HTML}
# https://cran.r-project.org/web/packages/kableExtra/vignettes/awesome_table_in_html.html#Overview
# all_crashes %>% filter(CYCLFLAG == "Y") %>% group_by(year = year(CRSHDATE), URBRURAL) %>% summarise(count = n()) %>% pivot_wider(names_from = URBRURAL, values_from = count) %>% adorn_totals("col") %>% kbl(caption = "Title") %>%  kable_styling()
```
