---
title: "rest_api_analysis"
output: html_document
---
This script imports court case data from the API output. It is a flattened JSON with 2,000+ variable names. The purpose of this notebook is to find OWI offenders, find if they have a 3+ OWI offense, then find sentencing length under that specific citation. A flattened JSON files was a bit tricky to work with as a case may have multiple offenses (i.e. charges) under different branches. I wrote scripts that grabs the column name, chops off the end part, then grabs sentence length under that exact same branch to make sure I'm grabbing the correct sentencing data.

Functions include:
  + Gather all OWI cases by performing a text search of the statute_citations column
  + Then also finds if a 3+ offender by performing another text search
  + Gather sentencing data of certain case types (done either by a specific sentenceCode or conditionCode). Depending on which branch these codes are located, this grabs the associated sentence length data. Most court cases should have exactly 1 of these codes, but the code does account for both. Later on, the code is written so that condCode takes priority, if there is data in both areas.
  + I run these functions for each charge in a case as to look under different branches. The functions do have a charge_no variable. I found most cases have 12 or fewer charges
  + Then the dataframe of these 12 charges (that have sentencing data) are merged into 1 clean dataframe and then I did my analysis

```{r setup, include=FALSE}
# Set working directory to where this notebook is saved
knitr::opts_chunk$set(echo = FALSE, comment = NA, warning = FALSE, error = FALSE, message = FALSE, tidy = TRUE, fig.dim = c(7, 4), global.par = TRUE)
knitr::opts_knit$set(root.dir = '/...')
```

```{r libraries}
library(dplyr)
library(data.table)
```

```{r Load the data}
cases <- readRDS( "C:/REST_API_Searches/x.rds")
```

```{r Functions to grab OWI cases and Offense Numbers}
# 346.63(1)(a) and (1)(b) only want these, one could be charged with both

# Select all columns with statuteCite and filters ones that match our OWI statute,
get_OWI_cases <- function(dataframe) { # .*\\. acts as a wild card
  dataframe %>% select(id, matches("^charges.*\\.statuteCite$")) %>%
      filter(apply(., 1, function(thisrow) # need the \\ because of the parenthesis (regex uses ())
        any(grepl("^346.63\\(1\\)\\(a\\)|^346.63\\(1\\)\\(b\\)", thisrow) #^ means that beginning must match
        )))
}

# Searches for the third+ keywords, careful this could include non-OWI charges, make sure input has charges.x.descr specifically
get_third_plus_offenders <- function(dataframe) {
  dataframe %>% # select(id, matches("(^charges)(.descr$)")) %>%
      filter( apply(., 1, function(thisrow)
        any(grepl("3RD|3rd|4th|5th+|5th|6th|7th|8th|9th", thisrow) #^ means that beginning must match
        )))
}
```

```{r Function to get word from length abbreviations}
# Use in conjunction with mutate to add a column to the df, for pdCode
get_pdCode_descr <- function(df,colname) {
  case_when(
    colname == "DA" ~ "Days",
    colname == "HR" ~ "Hours",
    colname == "LF" ~ "Life",
    colname == "MN" ~ "Minutes",
    colname == "MO" ~ "Months",
    colname == "WK" ~ "Weeks",
    colname == "YR" ~ "Years",
    colname == colname ~ colname
  )
}
```

```{r Function to get Sentencing by SentCode}
get_sentence_sentCode <- function(df, chargecol) {
  # sentCodes of State prison, Local jail and House of Correction (MKE only)
  # sentcodes = c("73", "74", "90") # 3 codes, that's why there's 3 columns
  codes <- # Select id and sentcode columns only
    df %>% select_at(vars(id, grep(
      paste0("^", chargecol, ".*\\.sentCode$"), colnames(.)
    )))
  # Create 3 columns that have the match, puts respected colname as the variable so we know actually where to grab sentencing info bc this is flattened JSON
  # cant do codes == c("73", "74", "90") for some reason??
  codes$colname1 = apply(
    codes == "73",
    1,
    FUN = function(x)
      toString(names(x)[x])
  )
  codes$colname2 = apply(
    codes == "74",
    1,
    FUN = function(x)
      toString(names(x)[x])
  )
  codes$colname3 = apply(
    codes == "90",
    1,
    FUN = function(x)
      toString(names(x)[x])
  )

  # Remove crap so we only get the text
  # There could be more than 2 matches, if so these will be NA, not included, not sure how else to do this
  codes$colname1 <-
    gsub(",", "", codes$colname1)
  codes$colname1 <-
    gsub("NA", "", codes$colname1)
  codes$colname1 <- trimws(codes$colname1)
  
  codes$colname2 <-
    gsub(",", "", codes$colname2)
  codes$colname2 <-
    gsub("NA", "", codes$colname2)
  codes$colname2 <- trimws(codes$colname2)
  
  codes$colname3 <-
    gsub(",", "", codes$colname3)
  codes$colname3 <-
    gsub("NA", "", codes$colname3)
  codes$colname3 <- trimws(codes$colname3)

  # May accidentally match with id, remove when variable is 'id'
  codes <- data.frame(lapply(codes, function(x) {gsub("id", "", x)})) %>% mutate(id = as.integer(id))
  
  # Counts how repeats there will be due to having matching sentCode. Later on these matches will be merge into just one column
  codes_sel <- codes %>% select(colname1, colname2, colname3)
  
  codes_sel[(grepl("sentCode", codes_sel$colname1) & grepl("sentCode", codes_sel$colname2)) |(grepl("sentCode", codes_sel$colname1) & grepl("sentCode", codes_sel$colname3)) | (grepl("sentCode", codes_sel$colname2) & grepl("sentCode", codes_sel$colname3)), ] %>% nrow()
  
  # Count where there's a match - 1,322 for first charge out of 2026
  # codes %>% filter(colname1 != "" | colname2 != "" | colname3 != "") %>% nrow()
  
  # Replace sentCode column name with len1 so we grab the associated column names, do this for each colname. String MUST match column name we will select.
  codes <-
    codes %>% mutate(len1 = gsub("sentenceCode.sentCode", "len1", colname1))
  codes <-
    codes %>% mutate(time_period1 = gsub("sentenceCode.sentCode", "pd1.descr", colname1))
 
  codes <-
    codes %>% mutate(len2 = gsub("sentenceCode.sentCode", "len1", colname2))
  codes <-
    codes %>% mutate(time_period2 = gsub("sentenceCode.sentCode", "pd1.descr", colname2))
 
  codes <-
    codes %>% mutate(len3 = gsub("sentenceCode.sentCode", "len1", colname3))
  codes <-
    codes %>% mutate(time_period3 = gsub("sentenceCode.sentCode", "pd1.descr", colname3))
  
  # Combine back to original df
  combine <- codes %>% select(id, starts_with("len"), starts_with("time_period")) %>% left_join(df, codes, by = "id")
  
  # Loops through each row and grabs sentence time/length using the len/time_period columns data as a string (i.e. the colname). Do this 3 times for each column.
  combine$len_num1 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["len1"]]
  )
  combine$time_period_num1 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["time_period1"]]
  )
    combine$len_num2 = apply(
    combine, 
    1,
    FUN = function(x)
      x[x["len2"]]
  )
  combine$time_period_num2 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["time_period2"]]
  )
    combine$len_num3 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["len3"]]
  )
  combine$time_period_num3 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["time_period3"]]
  )
  # Returns only a few columns
combine %>% select(id, starts_with("time_period_num"), starts_with("len_num"))
  
  # 3518 has 2 codes of 74 ??
}
```

```{r Function to get Sentencing by condCode}
get_sentence_condCode <- function(df, chargecol) {
  # This is wcisCondCodes in the API condCode HC = House of Correction, A = Jail Time
  # condCodes = c("HC","A")
  # First, select the id and all condCode columns
  codes <-
    df %>% select_at(vars(id, grep(
      paste0("^", chargecol, ".*\\.condCode$"), colnames(.)
    )))
  # Create 3 columns that have the match, puts respected colname as the variable
  codes$colname1 = apply(
    codes == "HC",
    1,
    FUN = function(x)
      toString(names(x)[x])
  )
  codes$colname2 = apply(
    codes == "A",
    1,
    FUN = function(x)
      toString(names(x)[x])
  )
  # Remove crap so we only get the text
  codes$colname1 <-
    gsub(",", "", codes$colname1)
  codes$colname1 <-
    gsub("NA", "", codes$colname1)
  codes$colname1 <- trimws(codes$colname1)
  
  codes$colname2 <-
    gsub(",", "", codes$colname2)
  codes$colname2 <-
    gsub("NA", "", codes$colname2)
  codes$colname2 <- trimws(codes$colname2)

  # Replace sentCode column name with len1 so we grab the associated column names, do this for each colname
  codes <-
    codes %>% mutate(len1 = gsub("condition.condCode", "condLen1", colname1))
  codes <-
    codes %>% mutate(time_period1 = gsub("condition.condCode", "condPd1.pdCode", colname1))
 
  codes <-
    codes %>% mutate(len2 = gsub("condition.condCode", "condLen1", colname2))
  codes <-
    codes %>% mutate(time_period2 = gsub("condition.condCode", "condPd1.pdCode", colname2))

  # Combine back to original df
  combine <- codes %>% select(id, starts_with("len"), starts_with("time_period")) %>% left_join(df, codes, by = "id")
  
  # Loops through each row and grabs sentence time/length using the len columns data as a variable (i.e. the colname)
  combine$len_num1 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["len1"]]
  )
  combine$time_period_num1 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["time_period1"]]
  )
    combine$len_num2 = apply(
    combine, 
    1,
    FUN = function(x)
      x[x["len2"]]
  )
  combine$time_period_num2 = apply(
    combine,
    1,
    FUN = function(x)
      x[x["time_period2"]]
  )
  # Counts how many NA there will be due to having matching sentCode
  codes_sel <- codes %>% select(colname1, colname2)
  
  codes_sel[(grepl("sentCode", codes_sel$colname1) & grepl("sentCode", codes_sel$colname2)), ] %>% nrow()
  
  combine %>% mutate(time_period_num1 = get_pdCode_descr(., time_period_num1), time_period_num2 = get_pdCode_descr(., time_period_num2)) %>% select(id, starts_with("time_period_num"), starts_with("len_num"))
}
```

```{r Function to merge all sentences together}
# Merge sentences - this is used inside get_third_plus_offenders()
merge_sentence_length <- function(owi_data) {
  # Select time column created in if_owi_citation(), merges all of those columns which should result i only 1 variable
  sen_length <- owi_data %>% select(matches(c("^len_num\\d"))) # %>% select_if( ~ any(!is.na(.)))
  sen_length <- apply(sen_length, 1, function(x) x[!is.na(x)][1])
  
  # Same as above except for time period (Hours, Days, Months, Years)
  sen_time <- owi_data %>% select(matches(c("^time_period_num\\d"))) # %>% select_if( ~ any(!is.na(.)))
  sen_time <- apply(sen_time, 1, function(x) x[!is.na(x)][1])
  
  owi_data %>% mutate(length = sen_length,
                      time_length = sen_time)
}

# The last step of the analysis - make a tidy dataframe - separated by time.x and time.y for sentCode and condCode. Merges x and y, respectively
merge_sentence_length_xy <- function(owi_data) {
  # Select time column created in if_owi_citation(), merges all of those columns which should result i only 1 variable
  sen_length.x <- owi_data %>% select(matches(c("^time.x"))) # %>% select_if( ~ any(!is.na(.)))
  sen_length.x <- apply(sen_length.x, 1, function(x) x[!is.na(x)][1])
  
  # Same as above except for time period (Hours, Days, Months, Years)
  sen_time.x <- owi_data %>% select(matches(c("^time_per.x"))) # %>% select_if( ~ any(!is.na(.)))
  sen_time.x <- apply(sen_time.x, 1, function(x) x[!is.na(x)][1])
  
  sen_length.y <- owi_data %>% select(matches(c("^time.y"))) # %>% select_if( ~ any(!is.na(.)))
  sen_length.y <- apply(sen_length.y, 1, function(x) x[!is.na(x)][1])
  
  # Same as above except for time period (Hours, Days, Months, Years)
  sen_time.y <- owi_data %>% select(matches(c("^time_per.y"))) # %>% select_if( ~ any(!is.na(.)))
  sen_time.y <- apply(sen_time.y, 1, function(x) x[!is.na(x)][1])
  
  owi_data %>% mutate(length.x = sen_length.x,
                      time_length.x = sen_time.x,
                      length.y = sen_length.y,
                      time_length.y = sen_time.y)
}
```

```{r Function to find OWI, third+ offenders, and return sentence info}
# Find if citation matches OWI and returns new columns for sentence length and time period
# The function also filters if third+ offender - using get_third_plus_offenders()
# regex_charge is to find columns of that citation number
if_owi_citation <-
  function(casedata, # data
           charge_start, # regex for statutecitation for that particular charge no
           chargecol, # regex for all the 'Charges' columns to find that particular charge no
           number) {
    # charge no, used to make a unique column name
    
    # special case for chargeNo 1, do this so we can select charge.descr for statuteCite
    if (chargecol == "charges.[[:alpha:]]") {
      chargecol_special = "charges."
    } else {
      chargecol_special = chargecol
    }
    # Statutes to look for - if OWI
    pattern = "^346.63\\(1\\)\\(a\\)|^346.63\\(1\\)\\(b\\)"
    
    # Select columns that we need, not necessary
    charge <- casedata %>% select(
      id,
      #caseNo, county.countyNo,
      matches(charge_start),
      # statuteCite
      grep(paste0("^", chargecol, ".*\\.descr$"), names(.)),
      # all ending in descr,
      paste0(chargecol_special, "descr"),
      # charge description
      grep(paste0("^", chargecol, ".*\\.sentCode$"), names(.)),
      # sentCode
      grep(paste0("^", chargecol, ".*\\.condCode$"), names(.)),
      # condCode
      # matches("(^charges)(.descr$)"),
      # These last 4 are for sentence length (i.e. 18) and sentence time period (i.e. Months)
      grep(paste0("^", chargecol, ".*\\.condLen1$"), names(.)),
      grep(paste0("^", chargecol, ".*\\.condPd1.descr$"), names(.)),
      grep(paste0("^", chargecol, ".*\\.condPd1.pdCode$"), names(.)),
      
      grep(paste0("^", chargecol, ".*\\.len1$"), names(.)),
      grep(paste0("^", chargecol, ".*\\.pd1.descr$"), names(.))
    )
    
    # Find if statute is OWI statute
    # This iterates through every row and finds if OWI statute is a match,
    # returns id and statutes, so must late join back with charge time
    owi <-
      charge %>% select(id,
                        matches(charge_start),
                        paste0(chargecol_special, "descr")) %>%
      filter(apply(., 1, function(.)
        any(grepl(pattern, .))))
    # Also put find if third offender then join back to data in the Return
    third <- get_third_plus_offenders(owi)
    # Joins back, this time with filtered data
    charge_time <- semi_join(charge, third, by = "id")
    
    # If there is no sentCode, return NA
    # if (length(sentcode_colname) == 0) {
    if (nrow(charge_time) == 0) {
      return()
    } else {
      # This returns a df of sentence times and lengths using sentCode
      sentence_sentCode <-
        get_sentence_sentCode(charge_time, chargecol)
      
      # Use this function to combine all columns so only 1 variable remains of each
      sentence_sentCode_combine <-
        merge_sentence_length(sentence_sentCode) %>% select(id, length, time_length)
      
      # This returns a df of sentence times and lengths using condCode
      sentence_condCode <-
        get_sentence_condCode(charge_time, chargecol)
      
      # Use this function to combine all columns so only 1 variable remains of each
      sentence_condCode_combine <-
        merge_sentence_length(sentence_condCode) %>% select(id, length, time_length)
      # 1555 / 2026 have data - per codetype 1,307 (15 had multiple, were excluded) and 447 have a match
      # Combine these 2 so we have length.x (sentCode) and length.y (condCode)
      sentence_combine <-
        left_join(sentence_sentCode_combine,
                  sentence_condCode_combine,
                  by = "id")
      
      # Add this data back into the df
      charge_time <- left_join(charge_time, sentence_combine, by = "id")
      
      # Rename these columns so they're unique cause we'll join them later, old name, new name
      setnames(charge_time, "length.x", paste0("time.x", number))
      setnames(charge_time, "length.y", paste0("time.y", number))
      setnames(charge_time,
               "time_length.x",
               paste0("time_per.x", number))
      setnames(charge_time,
               "time_length.y",
               paste0("time_per.y", number))
      setnames(
        charge_time,
        paste0(chargecol_special, "descr"),
        paste0(chargecol_special, "descr", number)
      )
      charge_time
    }
  }

```

```{r Cases Analysis}
# Check statute numbers are consistent
cases$charges.statuteCite %>% unique() %>% sort()

# Find number of empty ones because of >12 citations
cases %>% filter(is.na(charges.statuteCite)) %>% nrow()

# Find duplicates
cases %>% group_by(county.countyNo, caseNo) %>% summarise(x=n()) %>% filter(x>1)

#Find count w/o duplicates just to make sure we get the right number
cases %>% distinct(county.countyNo, caseNo)

# Remove dupicates - MAKE SURE TO DO THIS!!!
cases <- unique(setDT(cases), by = c("county.countyNo", "caseNo"))

# Look at duplicated cases
# cases %>% filter(caseNo == "2019CM000093")#co 22 july 8
# get_case_data(df_row = get_casenumbers(b=baseurl, fdate = "2019-07-08", code = "28100")[18,], b = baseurl)
```
```{r Summarise our data}
# Get only OWI cases and join back to data
owi_cases <- get_OWI_cases(cases)
owi_cases <- semi_join(cases, owi_cases, by = "id")

# write.csv(owi_cases, "jan_dec_2019_owi_cases.csv")

# Look at charges descriptions. Search is grepl("3RD|3rd|4th|5th+|5th|6th|7th|8th|9th", thisrow), make sure this is right - .*\\. means wildcard
# This filters charges that are not already in grepl() above
owi_cases %>% select(matches(c("(^charges.*\\.descr$)"))) %>% tidyr::gather() %>% select(value) %>% unique() %>% filter(across(value, ~ !grepl('3RD|3rd|4th|5th+|5th|6th|7th|8th|9th', .)))
```

```{r Run to get all sentence lengths of third offenders }
# Do the analysis!
# Verify there are 11 total charges cases$chargex.x
# This creates 4 new columns - timex and time_perx
  a1 <- if_owi_citation(cases, "charges.statuteCite", "charges.[[:alpha:]]", "1")
  a2 <- if_owi_citation(cases, "charges.2.statuteCite", "charges.2.", "2")
  a3 <- if_owi_citation(cases, "charges.3.statuteCite", "charges.3.", "3")
  a4 <- if_owi_citation(cases, "charges.4.statuteCite", "charges.4.", "4")
  a5 <- if_owi_citation(cases, "charges.5.statuteCite", "charges.5.", "5")
  a6 <- if_owi_citation(cases, "charges.6.statuteCite", "charges.6.", "6")
  a7 <- if_owi_citation(cases, "charges.7.statuteCite", "charges.7.", "7")
  a8 <- if_owi_citation(cases, "charges.8.statuteCite", "charges.8.", "8")
  a9 <- if_owi_citation(cases, "charges.9.statuteCite", "charges.9.", "9")
  a10 <- if_owi_citation(cases, "charges0.statuteCite", "charges0.", "10")
  a11 <- if_owi_citation(cases, "charges1.statuteCite", "charges1.", "11")
  a12 <- if_owi_citation(cases, "charges2.statuteCite", "charges2.", "12")

# Combine all non-empty df
all_owi_sentence <- Reduce(function(x, y) merge(x, y, all=TRUE, by = "id"), list(a1,a2,a3,a4,a5,a6,a7,a8)) %>% select_if(~sum(!is.na(.)) > 0)

# Next, make into a singe pretty df
all_lengths <- merge_sentence_length_xy(all_owi_sentence) %>% select(id, length.x, time_length.x, length.y, time_length.y)

# We are looking at both sentCode and condCode - Tyler says to use condCode first, if that is NA then use sentCode. x is sentCode and y is condCode
all_lengths <- all_lengths %>% mutate(length_final = ifelse(is.na(length.y), length.x, length.y),
                       time_length_final = ifelse(is.na(time_length.y), time_length.x, time_length.y))

# Find how many have a sentence time
all_lengths %>% filter(!is.na(time_length_final)) %>% nrow()

# Get average jail time, adds a columns that calculates days
all_lengths <- all_lengths %>% mutate(length_final = as.integer(length_final))
# Get average jail time, adds a columns that calculates days
all_lengths <- all_lengths %>% mutate(time_in_days = case_when(time_length_final == "Months" ~ length_final * 30.4,
                                                time_length_final == "Hours" ~ length_final / 24,
                                                time_length_final == "Days" ~ length_final+0,
                                                time_length_final == "Years" ~ length_final * 365))

all_lengths %>% filter(!is.na(time_in_days)) %>% summarise(mean(time_in_days))

# Number with 10 or fewer days
all_lengths %>% filter(time_in_days < 11) %>% nrow()

# Look up a case
cases %>% filter(id == 1046) %>% select(county.countyNo, caseNo)
```
```{r}
# Cases with no sentencing
no_sentence <- all_lengths %>% filter(is.na(length.x) & is.na(length.y)) %>% semi_join(cases, . , by = "id")
no_sentence %>% select(charges.statuteCite, charges.jdgmts.dispo.descr) %>% table() %>% knitr::kable()

no_sentence %>% filter(charges.plea.descr == "Guilty") %>% select(id)
no_sentence %>% filter(charges.jdgmts.supervisions.len1 >1) %>% select(id)
no_sentence %>% filter(charges.jdgmts.timeConds.condLen1 >1) %>% select(id)
```
```{r}
# Find all sentence info for the first charge. 6222
sentCode <- get_sentence_sentCode(owi_cases, chargecol = "charges.")
condCode <- get_sentence_condCode(owi_cases, chargecol = "charges.")
code_join <- inner_join(condCode, sentCode, by = "id")

# Number with multiple times for sent and cond
time <- code_join %>% select(starts_with("time_per")) %>% mutate(x = rowSums(is.na(.))) %>% filter(x < 4)

time <- code_join %>% select(starts_with("time_per") & ends_with("x")) %>% mutate(x = rowSums(is.na(.))) %>% filter(x == 0)

time <- code_join %>% select(starts_with("time_per") & ends_with("y"), time_period_num3) %>% mutate(x = rowSums(is.na(.))) %>% filter(x < 2)

all_owi_sentence
# Number with multiple times for sent and cond for all 3+ owi. y is condCode - use this first

# Cases with both codes
time <- all_owi_sentence %>% select(starts_with("time_per")) %>% mutate(x = rowSums(is.na(.))) %>% filter(x < 11)

time <- all_owi_sentence %>% select(starts_with("time_per") & contains("x")) %>% mutate(x = rowSums(is.na(.))) %>% filter(x < 5)

time <- all_owi_sentence %>% select(starts_with("time_per") & contains("y")) %>% mutate(x = rowSums(is.na(.))) %>% filter(x < 5)

# Cases with both citations
all_owi_sentence %>% filter(apply(., 1, function(thisrow) any(grepl(c("^346.63\\(1\\)\\(a\\)"), thisrow)) & any(grepl(c("^346.63\\(1\\)\\(b\\)"), thisrow))))
```
